---
title: "Vectorized Operations"
---

One thing that I have always struggled with, going from programming in Python to R, is vectorized operations. I feel like loops make a lot of sense, and when planning out the logic of some code, or writing it, I can easily figure out how to do it with loops, but I struggle figure out how to think about it with vectorized operations. I think part of that is that it just takes a mentality shift and the other part is that I don't know how the functions for vectorized operations work well enough. Therefore, with this project, I will take a deep-dive into vectorized operations and get some practice with it, so that I hopefully feel confident using vectorized operations later on when writing code.

Goals for the project:

* Learning about the logic of vectorized operations
* Learning about the different functions for vectorized operations in R, including apply(), lapply(), and mapply()
* Comparing and contrasting writing vectorized vs looped code
* Getting more experience with vectorized operations so that I can confidently apply it to future coding projects

<br>

# The Basics of Vectorized Operations

<br>

## Examples of simple vectorized operations

This [book chapter](https://bookdown.org/rdpeng/rprogdatascience/vectorized-operations.html) had a good example of really basic vectorized operations. Contrasting the following two ways of adding two vectors:

```{r vect basic ver 1}
x <- 1:4
y <- 6:9 
z <- x + y
z
```
```{r vect basic ver 2}
x <- 1:4
y <- 6:9
for(i in seq_along(x)) {
      z[i] <- x[i] + y[i]
}
z
```

When seeing it written out like that, I realize that I already use some basics of vectorized operations.

Similarly, this is also vectorized operations:

```{r vect operation example}
x <- 1:4

x > 2
```
This is with matrices:

```{r vect operation simple matrix}
x <- matrix(1:4, 2, 2)
y <- matrix(rep(10, 4), 2, 2)
print(x)
print(y)
x * y 
```

<br>

## The different functions for vectorized operations in r

This information is taken from  on chapter 16 - [Loop Functions](https://bookdown.org/rdpeng/rprogdatascience/loop-functions.html) in the open source textbook "R programming for data science"

<br>

### lapply()

lapply() returns a list, hence the "l".

It takes three arguments, a list to do an operation on, the function, and any other parameters

If the input is not a list, it will be converted to a list automatically using as.list()

When you pass a function through lapply() (second parameter), it will input the list you have as the first parameter of that function. For example, in mean(), the first argument is the object to calculate the mean from and the rest of the parameters include e.g. whether parts of the object should be trimmed (with trim).

If the function you pass through lapply() has more parameters that you want to specify, such as trim = 0 in mean(), you can add those as extra parameters to lapply() and they will be passed down into the function you specify. For example, you could here have written lapply(x, mean, trim = 0).

You can enter functions that you define yourself into lapply().
You can even define it in the lapply() function itself. for example as "lapply(x, function(elt) { elt[,1] })". This creates an **anonymous function** that only exists for the lapply() command and then disappears.

<br>

### sapply()

sapply() is very similar to lapply(), but it simplifies the results. For example, if you run lapply() on a vector with three items, it will return a nested list where each list has one item each. However, if you run sapply() it will make it one vector with several items.
Similarly, if all lists that you input has the same number of items, sapply() will make it into a matrix, instead of nested lists. However, if none of these two situations apply, it will just produce the same nested lists as lapply() does.

See here for an example taken from the book chapter:

```{r lapply nested list example}
x <- list(a = 1:4, b = rnorm(10), c = rnorm(20, 1), d = rnorm(100, 5))
lapply(x, mean)
```

```{r sapply list example}
x <- list(a = 1:4, b = rnorm(10), c = rnorm(20, 1), d = rnorm(100, 5))
sapply(x, mean)
```

<br>

### vapply()

Similar to sapply and lapply(), but you can define what type of input you are expecting. For example, using FUN.VALUE = numeric(1) says that you are expecting the input to be numeric, with a single digit. FUN.VALUE = data.frame(1) indicates that you are expecting a single data frame. If the input is not in the excepted format, you will get an error. This can be helpful for example if you have string numbers, but want them to be numeric. Using vapply() instead of sapply() or lapply() then adds in a check for variable type and can create an error, instead of having a silent error happen later in the code. There is an argument for simplifying the output that produces similar output as lapply() or sapply().

<br>

### tapply()

It sort of groups part of your vector together. So instead of maybe having a nested list like x <- c((1, 2, 3,),(4, 5, 6),(7, 8, 9)), you can have the list x <- c(1, 2, 3, 4, 5, 6, 7, 8, 9) and the list y <- c(1, 1, 1, 2, 2, 2, 3, 3, 3). You can then define to group the calculations in x by the membership in y with tapply().

For example, tapply(x, y, mean)

There is also a parameter called simplify (TRUE or FALSE). If it is true it gives output similar to that of sapply(), while if it is false, it will give output similar to that of lapply(). If it is true, the output will always be an array. An array is similar to a matrix, as a matrix is a 2 dimensional array. 

<br>

### apply()

Used on an array, like a matrix. You can specify which direction you want it to act on, either rows or columns.

The second parameter, called MARGINS, determine whether you apply the function to the column or the row. If the value is 1, you collapse the columns and retain the rows. If the value is 2, you collapse the rows and retain the columns. See the following examples, taken from the bookchapter, for reference:

```{r apply code example column}
x <- matrix(rnorm(200), 20, 10)
apply(x, 2, mean)  ## Take the mean of each column
```

```{r apply code example row}
x <- matrix(rnorm(200), 20, 10)
apply(x, 1, sum)   ## Take the mean of each row
```
The following are some examples of functions that could be written as apply()

rowSums = apply(x, 1, sum)
rowMeans = apply(x, 1, mean)
colSums = apply(x, 2, sum)
colMeans = apply(x, 2, mean

<br>

### mapply()

mapply() can iterate over multiple r objects at the time, like running lapply() in paralell.

For mapply(), the first parameter is the function, not the object to run it for.


Here is an example of how mapply() saves space. For example, instead of writing

```{r no mapply example}
list(rep(1, 4), rep(2, 3), rep(3, 2), rep(4, 1))
```

You can use mapply() and write

```{r mapply example}
mapply(rep, 1:4, 4:1)
```

Here, 1:4 is passed to rep as the first parameter and 4:1 as the second.

You can also use mapply() to run vectorized operations with functions that do not support vectorized operations. For example, here is how the book chapter demonstrated it:

```{r vect mapply def function}
sumsq <- function(mu, sigma, x) {
         sum(((x - mu) / sigma)^2)
}

# Generate data in a list
x <- rnorm(100)
```

Here we are defining a function to calculate sums of squares and we want to do it for values of mu and sigma increasing from 1 to 10. If we try doing this in the function itself, which is not adapted for vectorized operations, we only get one number.

```{r vect mapply wrong example}
sumsq(1:10, 1:10, x)  ## This is not what we want
```
While we can apply mapply() to run the 10 values in paralell.

```{r vect mapply right example}
mapply(sumsq, 1:10, 1:10, MoreArgs = list(x = x))
```

You can also apparently do this with the Vectorize() function

```{r Vectorized function example}
vsumsq <- Vectorize(sumsq, c("mu", "sigma"))
vsumsq(1:10, 1:10, x)
```

I don't think I fully understand how this works yet, but I might get it a little more later. Or it might not be the most important thing to know.

mapply differs from e.g. lapply if you have multiple lists and elements as lapply iterates through each list in total, while mapply iterates through each element in a list, not necessarily iterating over each list per iteration like lapply does. See exercise 13 for an example.

<br>

### split()

The book chapter also mentions split(), as this seems to often be used with vectorized operations. You can use it to divide out a part of the items and run a function only on these. It takes the parameters x (vector or data frame), function, and drop = TRUE/FALSE. Seems like it works with factors.

```{r split funct example}
x <- c(rnorm(10), runif(10), rnorm(10, 1))
f <- gl(3, 10)
print(f)
split(x, f)
```

This seems very similar to tapply(), but with factors instead of just a list.

This is an example in combination with lapply()

```{r split funct and lapply}
lapply(split(x, f), mean)
```

With this, you can for example look at a data frame and split up the calculations or stats based on a variable (as a factor), such as month or city. You can even split it up in combinations of two or more variables, e.g. with (split(x, list(f1, f2))).

Remember from earlier, that tapply() is like split() and lapply(), but for vectors only. That means that you cannot use it on data frames.

<br>

## Check-in

Ok, I think I am starting to get a basic overview of the different functions, what they do, and how they work. Now, I think I might need some practice actually using them. I have asked ChatGPT to come up with exercises for me, so I will use that to get more comfortable using these vectorized operations.

<br>

# Practice!

So far, I have learned about apply(), lapply(), sapply(), vapply(), tapply(), mapply(), and split() with lapply(). I know need some practice using these functions to get comfortable with the logic behind them, when to use each, and how they work.

First, I fed ChatGPT with all the notes I have up until this point and asked it to make some exercises, with varying difficulty, that could be used to practice what is in this document so far and further deepen my understanding. Here are some exercises and my solutions to them from ChatGPT. Some of them might be slightly modified:

<br>

### 1. Simple Vectorized Operations
1.1. Create two vectors, a <- 1:5 and b <- 6:10. Then use vectorized operations for the next steps.

```{r ex 1.1.}
a <- c(1:5)
b <- c(6:10)
print(a)
print(b)
```

Little reminder to myself: lists differ from vectors as a vector can only have one type of items, e.g. numbers, characters, or logical values, while lists can have a mix. You define a vector with c() and a list with list(). I am pretty sure a c() with multiple data types will just automatically be converted to a list, since R tends to to that type of stuff.

1.2. Add, subtract, multiply, and divide the vectors.
```{r ex 1.2.}
c <- a + b
print(c)
d <- a - b
print(d)
e <- a * b
print(e)
f <- a / b
print(f)
```

As we learned about earlier, these kind of operations of the vectors is already vectorized, you don't have to use any of the functions like apply()


1.3. Find which elements of a are greater than 3.
```{r ex 1.3.}
g <- a > 3
print(g)
```

1.4. Rewrite the operations using loops instead of vectorized operations. I'll just do the last one from 1.3.
```{r ex 1.4.}
g <- c()
for (i in a) {
  if (i > 3) {
    g <- append(g, TRUE)
  } else {
    g <- append(g, FALSE)
  }
}
print(g)
```

That was indeed a lot more work than using vectorized operations.

<br>

### 2. Matrix Operations

2.1. Create a matrix mat <- matrix(1:9, nrow = 3).
```{r ex 2.1.}
mat <- matrix(1:9, nrow = 3)
print(mat)
```

2.2. Multiply each element of the matrix by 2 using vectorized operations.
```{r ex 2.2.}
mat_mult <- mat*2
mat_mult
```

2.3. Find the square of each element in the matrix.
```{r ex. 2.3.}
mat_square <- mat^2
mat_square
```

2.4. Check which elements are greater than 5.
```{r ex 2.4.}
mat > 5
```

<br>

### 3. apply()

Using apply()

3.1. Create a matrix mat <- matrix(rnorm(20), nrow = 5).
```{r ex 3.1.}
mat <- matrix(rnorm(20), nrow = 5)
mat
```

3.2. Use apply() to calculate the mean of each row.
```{r ex 3.2.}
apply(mat, MARGIN = 1,  mean)
```

3.3. Use apply() to find the sum of each column.
```{r ex 3.3.}
apply(mat, MARGIN = 2, sum)
```

3.4. Write a custom function that calculates the range (max - min) and use apply() to calculate the range for each row.
```{r ex 3.4.}
apply(mat, MARGIN = 1, function(x) {max(x)-min(x)})
```

I might actually be a genius. Just saying.

<br>

### 4. Using lapply() and sapply()

4.1. Create a list of numeric vectors:
```{r ex 4.1.}
my_list <- list(a = 1:5, b = rnorm(10), c = runif(20))
my_list
```

4.2. Use lapply() to calculate the mean of each vector.
```{r ex 4.2.}
lapply(my_list, mean)
```

4.3. Use sapply() to calculate the standard deviation of each vector.
```{r ex 4.3.}
sapply(my_list, sd)
```

4.4. Create an anonymous function inside lapply() to find the sum of squares for each vector.
```{r ex 4.4.}
lapply(my_list, function(x) {sum(x^2)})
```

<br>

### 5. Using tapply()

5.1. Create two vectors:
```{r ex 5.1.}
values <- c(10, 20, 30, 40, 50, 60, 70, 80)
groups <- c("A", "A", "B", "B", "C", "C", "A", "B")
```

5.2. Use tapply() to calculate the mean of values grouped by groups.
```{r ex 5.2.}
tapply(values, groups, mean)
```

5.3. Use tapply() to count the number of elements in each group.
```{r ex 5.3.}
tapply(values, groups, length)
```

<br>

### 6. Using mapply() - Advanced


6.1. Write a function power_fun <- function(base, exp) { base ^ exp }.
```{r ex 6.1.}
power_fun <- function(base, exp) { base ^ exp }
```

6.2. Use mapply() to apply this function to base = c(2, 3, 4) and exp = c(3, 2, 1).
```{r ex 6.2.}
base = c(2, 3, 4)
exp = c(3, 2, 1)
mapply(power_fun, base, exp)
```

6.3. Compare the output with the manual calculation of each base raised to its corresponding exponent.
```{r ex 6.3.}
for (i in 1:length(base)){
  print(power_fun(base[i],exp[i]))
}
```

<br>

### 7. Combining split() and lapply() - Advanced

7.1. Create a data frame:
```{r ex 7.1.}
df <- data.frame(
  id = 1:12,
  group = rep(c("X", "Y", "Z"), each = 4),
  score = rnorm(12, mean = 75, sd = 10)
)
df
```

7.2. Use split() to split the data frame by group.
```{r ex 7.2.}
split(df, df$group)
```

7.3. Use lapply() to calculate the mean score for each group.
```{r ex 7.3.}
lapply(split(df$score, df$group), mean)
```

7.4. Write a combined split() and lapply() operation to count the number of rows in each group.
```{r ex 7.4.}
lapply(split(df, df$group), nrow)
```

<br>

### 8. Simulating Data with mapply() - Advanced

8.1. Simulate data for 5 different groups, each with a mean and standard deviation:
```{r ex 8.1.}
means <- c(5, 10, 15, 20, 25)
sds <- c(1, 2, 3, 4, 5)
n <- c(10, 15, 20, 25, 30)
```

8.2. Use mapply() to generate random normal data for each group using rnorm().
```{r ex 8.2.}
mapply(rnorm, n, means, sds)
```

<br>

### 9. Combining Multiple Functions - Challenge

9.1. Create a data frame:
```{r ex 9.1.}
df <- data.frame(
  category = rep(c("A", "B", "C"), each = 5),
  value1 = rnorm(15, 50, 10),
  value2 = rnorm(15, 30, 5)
)
df
```

9.2. Use split() and lapply() to calculate the mean of value1 and value2 for each category.
```{r ex 9.2.}
lapply(split(df$value1, df$category), mean)
lapply(split(df$value2, df$category), mean)
```

9.3. Use mapply() to calculate the sum of squares for value1 and value2 for each category.
```{r ex 9.3.}
mapply(function(x) { sum(x)^2 }, split(df$value1, df$category))
mapply(function(x) { sum(x)^2 }, split(df$value2, df$category))

```

9.4. Compare the results with a loop-based approach.
```{r ex 9.4.}
vect_a <- c()
vect_b <- c()
vect_c <- c()
for (i in 1:length(df$category)){
  if (df$category[i] == "A") {
    vect_a <- append(vect_a, df$value1[i])
  } else if (df$category[i] == "B") {
    vect_b <- append(vect_b, df$value1[i])
  } else {
    vect_c <- append(vect_c, df$value1[i])
  }
}
(mean(vect_a))
(mean(vect_b))
(mean(vect_c))


```

That was a lot more extensive and thought-consuming than using vectorized operations, I see the appeal

<br>

### 10. Advanced Matrix Manipulations - Challenge

10.1 Create a 3D array:
```{r ex 10.1.}
arr <- array(1:24, dim = c(4, 3, 2))
arr
```

10.2. Use apply() to calculate the sum along each dimension.
```{r ex 10.2.}
apply(arr, MARGIN = 1, sum)
apply(arr, MARGIN = 2, sum)
apply(arr, MARGIN = 3, sum)
```

10.3. Use mapply() to scale the values in the array by different factors for each layer.
```{r ex 10.3.}
vect_mar <- c(1, 2, 3)
vect_scale <- c(10, 2, 3)
#mapply(function(x, y) {apply(arr, MARGIN = y, x*y)}, vect_mar, vect_scale)
```

I got stuck here. I don't know how to use anonymous functions to do this. This is the solution that ChatGPT gave (with different variable names):
```{rex 10.3. Chat}
arr <- array(1:24, dim = c(4, 3, 2))  # A 4x3x2 array
factors <- c(2, 3)  # Scaling factors for each layer

scaled_arr <- mapply(function(slice, factor) slice * factor,
                     split(arr, slice.index(arr, 3)),  # Split by layers
                     factors,
                     SIMPLIFY = FALSE)  # Avoid simplifying to vectors
scaled_arr <- array(unlist(scaled_arr), dim = dim(arr))  # Reshape to original dimensions
```

Turns out that I misunderstood what ChatGPT meant with layers. I though of that as being each dimension, but it meant each value of the third dimension. Here is the code ChatGPT gave me when I asked it to edit my code so that it would work:

```{r ex 10.3. Chat based on my code}
vect_mar <- c(1, 2)
vect_scale <- c(10, 2)

scaled_arr <- mapply(
  function(margin, scale) {
    # Giving the data set as all rows, all columns, but varying levels
    apply(arr[, , margin], MARGIN = c(1, 2), function(x) x * scale)
  },
  vect_mar,
  vect_scale,
  # This part makes it so that the matrices are kept separate in the output and not combined + somehow making the third column part of first and second
  SIMPLIFY = FALSE
)

scaled_arr
```

I sort of get it, but not 100%. I think my main challenge here though is the limitation in what I know about functions and maybe not so much the vectorized operations. Functions might be an interesting portfolio project for later. In the meantime, I will continue practicing vectorized operations.


Trying it again as of 1st of May, 2025:

10.3. Use mapply() to scale the values in the array by different factors for each layer.
```{r ex 10.3. later}
arr <- array(1:24, dim = c(4, 3, 2))
scale <- c(10, 2)
mapply(function(i, scale) {arr[,,i]*scale}, i = 1:2, scale, SIMPLIFY = FALSE)
```

I got it with a little help from ChatGPT. I was trying to do mapply(function(arr, i, scale) {arr[,,i]*scale}, arr, i = 1:2, scale, SIMPLIFY = FALSE), but it didn't work because it couldn't iterate through arr. So arr can be called in the anonymous function, but not as a parameter of it.

<br>

## Working from loops to vectorized operations - Exercises

The following section will have some exercises converting code using loops to vectorized operations. I wanted to do this because I feel like I can fairly easily think in terms of loops when designing code and I want to practice thinking with vectorized operations instead. In addition, it will help compare and contrast loops with vectorized operations. The exercises are taken from ChatGPT.

<br>

### 11. Multiply Rows of a Matrix by a Vector

Code with Loops:
```{r ex 11 original}
# A 4x3 matrix
mat <- matrix(1:12, nrow = 4, ncol = 3)

# A vector of scaling factors
scaling <- c(2, 4, 6, 8)

# Using a loop to multiply each row by its corresponding factor
result <- matrix(0, nrow = 4, ncol = 3)
for (i in 1:nrow(mat)) {
  result[i, ] <- mat[i, ] * scaling[i]
}
result
```

```{r ex 11 new}
mat <- matrix(1:12, nrow = 4, ncol = 3)
scaling <- c(2, 4, 6, 8)
#mapply(function(dat, scale) {})
```

I think, at this point, that I can't fully take advantage of the vectorized operations before I learn to deal with functions better. I will come back to this after I have worked through the module on functions. If I still feel like I need more practice after that module, I might make a separate portfolio project for that.

1st of May, 2025:
I am back, now at the end of the semester. Let us see whether I can do this exercise now.

```{r ex 11 new later}
mat <- matrix(1:12, nrow = 4, ncol = 3)
scaling <- c(2, 4, 6, 8)
scaled_mat <- mapply(function(mat, scaling) {mat*scaling}, mat, scaling)
scaled_mat <- matrix(scaled_mat, nrow = 4, ncol = 3)
scaled_mat
```

<br>

### 12. Calculate Column Means of a Matrix

Code with loops:
```{r ex 12 original}
# A 4x3 matrix
mat <- matrix(1:12, nrow = 4, ncol = 3)

# Using a loop to calculate the mean of each column
col_means <- numeric(ncol(mat))
for (j in 1:ncol(mat)) {
  col_means[j] <- mean(mat[, j])
}
col_means
```

Code rewritten with vectorized functions:
```{r ex 12 new}
# A 4x3 matrix
mat <- matrix(1:12, nrow = 4, ncol = 3)

apply(mat, 2, mean)
```

<br>

### 13. Scale Elements in a List

Code with loop:
```{r ex 13 original}
# A list of numeric vectors
lst <- list(a = 1:4, b = 5:8, c = 9:12)

# A vector of scaling factors
scaling <- c(2, 3, 4)

# Using a loop to scale each element of the list by its corresponding factor
scaled_lst <- list()
for (i in seq_along(lst)) {
  scaled_lst[[i]] <- lst[[i]] * scaling[i]
}
scaled_lst
```

Code with vectorized operations:
```{r ex 13 new}
# A list of numeric vectors
lst <- list(a = 1:4, b = 5:8, c = 9:12)

# A vector of scaling factors
scaling <- c(2, 3, 4)

#lapply(lst, function(lst, scaling) {lst*scaling}, scaling)

mapply(function(lst, scaling) {lst*scaling}, lst, scaling, SIMPLIFY = FALSE)
```

I first tried using lapply, which didn't work as it was going element by element instead of list by list. mapply works by iterating over parameters instead of elements. mapply therefore applies the nested lists pairwise (e.g. lst[1] with mat[1] and lst[2] with mat[2]). The functions therefore produce different results.

<br>

### 14. Calculate Grouped Means

Code with loops:
```{r ex 14 original}
# A numeric vector
x <- c(5, 10, 15, 20, 25, 30, 35, 40, 45)

# A grouping factor
group <- c(1, 1, 2, 2, 2, 3, 3, 3, 3)

# Using a loop to calculate the mean for each group
unique_groups <- unique(group)
group_means <- numeric(length(unique_groups))
for (i in seq_along(unique_groups)) {
  group_means[i] <- mean(x[group == unique_groups[i]])
}
group_means
```

Code with vectorized operations:
```{r ex 14 new}
# A numeric vector
x <- c(5, 10, 15, 20, 25, 30, 35, 40, 45)

# A grouping factor
group <- c(1, 1, 2, 2, 2, 3, 3, 3, 3)

tapply(x, group, mean)
```

Note that the output is a little different when using tapply(), as it shows the grouping number as well as the means, while the solution with loops only showed the means and the index alone identified the grouping variable.

<br>

### Reflection

I have no worked through several exercises using vectorized functions, and I think I am starting to get the logic behind it. I would still need some more practice before I could confidently use all of the functions, know their parameters, and the nuances between their outputs. However, I think I have achieved my goal of learning the logic of vectorized operations and being able to use them (with a little help from the help section).

<br>

## Anonymous Functions

I have used anonymous functions quite a bit throughout this project, so I thought I would formalize it by adding a section on what anonymous functions are and how to use them.

An anonymous function is a function that is not stored as a variable, but instead just used in that one instance. It therefore does not have a name (hence "anonymous"). The syntax for the anonymous functions is function(parameters) function_body. So you start with the word "function", then add any parameters you need, a space, and the action of the function. Anonymous functions are great for short functions, but can get messy if they are longer.

<br>

## Caveats to Vectorized Operations - Reflection after the end of the semester

After having learned more about R, worked on bigger projects in R, and learned more about Tidyverse, I have learned that altough the vectorized operations used here can often speed up the coding process and make it more readable, there are situations in which other approaches work better.

<br>

### Loops

First, as stated in the introduction of this project, I started out being very familiar with loops, but less so with vectorized operations. I have however, through working on some of the applied projects coming later, learned that loops sometimes work better than vectorized operations. This occurs in scenarios where one wants to use the outcome of one iteration in the next iteration. There might be some workarounds for doing this with vectorized operations, but it seems like loops (although a little slower) are easier to use for these purposes. However, in any other case, vectorized operations make clearer code and run faster than loops. Therefore, it seems like vectorized operations should usually be used when I think about using loops.

<br>

### Tidyverse solutions

Tidyverse has a lot of useful functions that can sometimes mitigate the need for using the vectorized functions investigated in the current project. Therefore, it is usually good practice to google whether there is a Tidyverse (or other) function for what you want to do before starting to write code for it yourself. I have found several functions I didn't know about earlier from googling when I have needed something and it has saved me from writing out a lot more complicated code. In addition, there are often easier Tidyverse solutions compared to writing code using the base R vectorized operations such as apply(). See next project for more information about Tidyverse vs base R.

<br>

## Cheat sheet

Here is a quick overview of the different functions in base R for vectorized operations and what they do, for reference later:

- lapply() - acts on a list and returns a list
  - Coerces the input to a list (nested if it has nested lists)
  - Has 3 parameters: list, function, parameters of the second function
- sapply() - like lapply(), but with simplified results
  - Results in unnested lists or matrices if all of the lists have the same lenght
    - If not, it produces the same results as lapply()
- vapply() - similar to lapply() and sapply(), but also validates the input type and length
  - Has a simplification parameter (similar to lapply() vs sapply())
- tapply() - applies a function to defined groupings in the list
  - Has three parameters: x-vector, y-grouping, and function
  - Has a simplification parameter (similar to lapply() vs sapply())
apply() - applied to an array, like a data frame or matrix
  - If the parameter margins is 1, it will apply the function to each row, while margins = 2 will apply the function to each column
  - The output will be a vector
- mapply() - Like running lapply() in parallel across parameters of a function
  - mapply(function, funct_param1, funct_param2)
- Combining split() and lapply() - like tapply(), but with factors instead of a grouping vector
  - Can also be used on data frames