---
title: "ggplot"
---

# General information about ggplot

Taken from module 2 and independent research

```{r libraries for demonstrations}
library(palmerpenguins)
library(ggplot2)
```

The ggplot function, in its most basic form, takes a dataset and the dimensions/axes of the plot. Then you add the actual data points with e.g. data_point()
This can for example look like ggplot(data = dataset, mapping = aes(x = x_var, y = y_var)) +
geom_point()
Mapping is to tell it what to map out on the plot, using aes to define what to put on it. If you just use mapping and not geom_point(), you will just get the axes on the plot, but no content. You therefore have to add that by e.g. geom_point().

Here is an example of what happens if you only run the general ggplot function, without e.g. geom_point()

```{r demonstration no geom_point}
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm, y = bill_length_mm,
                     color = species))
```
If you add geom_point(), you will get:


```{r demonstration with geom_point}
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm, y = bill_length_mm,
                     color = species)) +
  geom_point()
```


### Labels

Here is an example of different labels you can add to a graph and where it shows up:

```{r labels}
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm, y = bill_length_mm,
                     color = species)) +
  geom_point() +
  labs(title = "Bill depth and length",
       subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
       x = "Bill depth (mm)", y = "Bill length (mm)",
       color = "Species",
       caption = "Source: Palmer Station LTER / palmerpenguins package")
```


## Diff based on variable

We know that we can have variables distinguished by mapping them onto the x and y axes, as well as separate them by color. However, you can also separate them by shape, size, and alpha (transparency). All of these are done in aes(). You can also have one variable have two different characteristics, such as both different colors and shapes, to make the graph extra clear.

If you add these properties inside of e.g. the geom_point() instead of mapping and aes, you will just apply that to all data points, instead of have it differ based on a variable.

You can also facet your plot and get individual plot per combination of variable(s). If you have one, use facet_wrap(), if you have multiple, use facet_grid().
facet_grid(var_1 ~ var_2) or facet_wrap(~ var_1)
If you want facet_wrap(), but with max two columns, you can specify that, e.g. facet_wrap(~ var_1, ncol = 2)


## Random notes

If you want to remove the legend, add guides(color = FALSE) where you switch out color with whatever you used to separate the variables with. Guides is added as a separate function with the + sign to the general ggplot code.


