---
title: "ggplot and Data Visualization"
---

# General information about ggplot

Taken from module 2 and independent research

```{r libraries for demonstrations}
library(palmerpenguins)
library(ggplot2)
```

The ggplot function, in its most basic form, takes a dataset and the dimensions/axes of the plot. Then you add the actual data points with e.g. data_point()
This can for example look like ggplot(data = dataset, mapping = aes(x = x_var, y = y_var)) +
geom_point()
Mapping is to tell it what to map out on the plot, using aes to define what to put on it. If you just use mapping and not geom_point(), you will just get the axes on the plot, but no content. You therefore have to add that by e.g. geom_point().

Here is an example of what happens if you only run the general ggplot function, without e.g. geom_point()

```{r demonstration no geom_point}
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm, y = bill_length_mm,
                     color = species))
```
If you add geom_point(), you will get:


```{r demonstration with geom_point}
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm, y = bill_length_mm,
                     color = species)) +
  geom_point()
```


### Labels

Here is an example of different labels you can add to a graph and where it shows up:

```{r labels}
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm, y = bill_length_mm,
                     color = species)) +
  geom_point() +
  labs(title = "Bill depth and length",
       subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
       x = "Bill depth (mm)", y = "Bill length (mm)",
       color = "Species",
       caption = "Source: Palmer Station LTER / palmerpenguins package")
```

You can also add geom_text() or geom_label() (creates a box around the label) to get text on the plot itself. If you don't want all the labels for the data points, filter out for example the top 10 and plot those
  If your labels overlap, the easiest way to undo that is to use geom_labels_repel() in the ggrepel package


## Diff based on variable

We know that we can have variables distinguished by mapping them onto the x and y axes, as well as separate them by color. However, you can also separate them by shape, size, and alpha (transparency). All of these are done in aes(). You can also have one variable have two different characteristics, such as both different colors and shapes, to make the graph extra clear.

If you add these properties inside of e.g. the geom_point() instead of mapping and aes, you will just apply that to all data points, instead of have it differ based on a variable.

You can also facet your plot and get individual plot per combination of variable(s). If you have one, use facet_wrap(), if you have multiple, use facet_grid().
facet_grid(var_1 ~ var_2) or facet_wrap(~ var_1)
If you want facet_wrap(), but with max two columns, you can specify that, e.g. facet_wrap(~ var_1, ncol = 2)


## Random notes

If you want to remove the legend, add guides(color = FALSE) where you switch out color with whatever you used to separate the variables with. Guides is added as a separate function with the + sign to the general ggplot code.

If you have most of your data clustered around a smaller range of the plot. You can use filter and put a subsetted part of the data set into the plot, to see the nuances more clearly


## Different types of plots

### Histograms

```{r histogram}
ggplot(penguins, aes(x = flipper_length_mm)) +
  geom_histogram()
```

You can change the binwidth (the amount included in one bar in the histogram) by specifying binwidth in the geom_histogram() function. See examples:



```{r histogram high binwidth}
ggplot(penguins, aes(x = flipper_length_mm)) +
  geom_histogram(binwidth = 20)
```


```{r histogram small binwidth}
ggplot(penguins, aes(x = flipper_length_mm)) +
  geom_histogram(binwidth = 1)
```

You can add colors to fill in the histogram based on another variable. In that case, you should also specify alpha (transparency) to see the value lines better.


```{r histogram color}
ggplot(penguins, aes(x = flipper_length_mm, fill = species)) +
  geom_histogram(alpha = 0.60)
```

If you add fill you get the bars filled in, if you add color, you get the outline of the bar colored.


### Density plot


```{r density plot}
ggplot(penguins, aes(x = flipper_length_mm, color = species)) +
  geom_density()
```

You can add color to get individual lines with color, fill to fill in the lines (like mountains), shape seemingly just adds several lines that are all black and look the same. Size gives you different sizes, but it looks really ugly, like something you drew in paint.

You can also adjust the density (smoothness) of the curve with adjust. See examples


```{r density plot low adjustment}
ggplot(penguins, aes(x = flipper_length_mm, color = species)) +
  geom_density(adjust = 0.5)
```


```{r density plot high adjustment}
ggplot(penguins, aes(x = flipper_length_mm, color = species)) +
  geom_density(adjust = 2)
```

I think adjust = 1 is the standard.


### Box plot

```{r box plot}
ggplot(penguins, aes(x = flipper_length_mm)) +
  geom_boxplot()
```

You can also add variables to the y-axis:

```{r box plot y-axis}
ggplot(penguins, aes(x = flipper_length_mm, y = species)) +
  geom_boxplot()
```

### Scatterplot

```{r scatterplot}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
```


```{r scatterplot smooth}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  geom_smooth()
```


### Hex plot


```{r hex plot}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_hex()
```

### Bar chart

```{r bar chart}
ggplot(data = penguins, mapping = aes(x = species)) +
  geom_bar()
```

You can flip the plot on its side with coord_flip(). You can also technically do this by making the variable of interest the y variable instead of x variable.


```{r bar chart side-ways}
ggplot(data = penguins, mapping = aes(x = species)) +
  geom_bar() +
  coord_flip()
```

If you want to color in the bars by another variable, you can add fill = variable to color it in based on counts. You can also add in position = variable in geom_bar() to fill it based on its proportion (in addition to having fill in the aesthetics). See below:

Count:


```{r bar chart count}
ggplot(data = penguins, mapping = aes(x = species, fill = sex)) +
  geom_bar()
```

Proportion:

```{r bar chart proportion}
ggplot(data = penguins, mapping = aes(x = species, fill = sex)) +
  geom_bar(position = "fill")
```

### Violin plots

```{r violin plot}
ggplot(penguins, aes(x = species, y = flipper_length_mm)) +
  geom_violin()
```

### Ridge plots

Requires ggridges library. Looks like several valleys laying on top of each other.

```{r ridge plot}
# library(ggridges)
# ggplot(loans, aes(x = loan_amount, y = grade, fill = grade, color = grade)) + 
 # geom_density_ridges(alpha = 0.5)
```




## Customizing the layout

You can remove the tick marks and labels with theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank())
    
See example with box plot:

```{r box plot no y axes labels}
ggplot(penguins, aes(x = flipper_length_mm)) +
  geom_boxplot() + 
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank())
```

Note that the plot now has no labels or values on the y-axis.

You can also change themes to change the layout of the plot. 


```{r box plot bw theme}
ggplot(penguins, aes(x = flipper_length_mm)) +
  geom_boxplot() + 
  theme_bw()
```

```{r box plot minimal theme}
ggplot(penguins, aes(x = flipper_length_mm)) +
  geom_boxplot() + 
  theme_minimal()
```

## Weird things that I can't explain, but are probably useful to know for debugging

If you have a color layer to a plot and add geom_smooth() it gives you a smooth line for each category in geom too. You can fix this by moving the color argument of eas for ggplot into the specific geom_point() function, such as geom_point(color = species). However, for some reason, you can also fix it by adding color = "black" as an argument to geom_smooth(). 
See the following three examples

```{r weird smooth 1}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species)) +
  geom_smooth()
```

```{r weird smooth 2}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point() +
  geom_smooth()
```


```{r weird smooth 3}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point() +
  geom_smooth(color = "blue")
```



## Geoms

geom_rect() draws a rectangle around a point of interest
geom_segment() draws an arrow to the point of interest



## Libraries

ggpattern creates cool patterns, but crashes R when it gets too complicated. For example, this code works without ggpattern, but not with ggpattern. Other graphs that didn't have facetting worked with ggpattern. I checked this on the lab computer too and it still crashes, so it isn't due to my computer not being strong enough.
```{r thesis-plot-diff-scores-ext-age, eval = FALSE}
avg_diff_score_df %>% 
   select(diff_sc_ext_fm_mean, diff_sc_ext_fm_sd, diff_sc_ext_fc_mean, diff_sc_ext_fc_sd, diff_sc_ext_mc_mean, diff_sc_ext_mc_sd, Age_cat) %>%
  pivot_longer(cols = -Age_cat, names_to = c("raters", ".value"), names_pattern = "[[:alpha:]]+_sc_ext_([[:alpha:]]+)_([[:alpha:]]+)") %>%
  mutate(
    raters = case_when(
      raters == "fm" ~ "Father - Mother",
      raters == "fc" ~ "Father - Child",
      raters == "mc" ~ "Mother - Child"
    )
  ) %>% 
  ggplot(aes(x = Age_cat, y = mean)) +
  geom_bar(stat = "identity", color = "black", fill = "white", width = 0.4) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.20) +
  facet_wrap(vars(raters), nrow = 3) +
  #ylim(-10, 5) +
  theme_classic() +
  theme(axis.text.x = element_text(colour = "black"), 
        axis.text.y = element_text(colour = "black"),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.y = element_line(),
        panel.grid.minor.y = element_line(),
        strip.background = element_blank()) +
  #coord_cartesian(expand = FALSE, xlim = c(0.5, 3.5), ylim = c(-10, 5)) +
  geom_hline(aes(yintercept = 0)) +
  labs(
    x = "Age Category",
    y = "Mean Difference Scores"
  )
```

Internalizing scores:
```{r thesis-plot-diff-scores-int-age, eval = FALSE}
avg_diff_score_df %>% 
   select(diff_sc_int_fm_mean, diff_sc_int_fm_sd, diff_sc_int_fc_mean, diff_sc_int_fc_sd, diff_sc_int_mc_mean, diff_sc_int_mc_sd, Age_cat) %>%
  pivot_longer(cols = -Age_cat, names_to = c("raters", ".value"), names_pattern = "[[:alpha:]]+_sc_int_([[:alpha:]]+)_([[:alpha:]]+)") %>%
  mutate(
    raters = case_when(
      raters == "fm" ~ "Father - Mother",
      raters == "fc" ~ "Father - Child",
      raters == "mc" ~ "Mother - Child"
    )
  ) %>% 
  ggplot(aes(x = Age_cat, y = mean)) +
  geom_bar(stat = "identity", width = 0.4) +
  geom_col_pattern(aes(pattern = "mean"), pattern = "stripe", pattern_colour =  "lightgrey", color = "black", fill = "white", width = 0.4) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.20) +
  facet_wrap(vars(raters), nrow = 3) +
  #ylim(-10, 5) +
  theme_classic() +
  theme(axis.text.x = element_text(colour = "black"), 
        axis.text.y = element_text(colour = "black"),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.y = element_line(),
        panel.grid.minor.y = element_line(),
        strip.background = element_blank()) +
  #coord_cartesian(expand = FALSE, xlim = c(0.5, 3.5), ylim = c(-10, 5)) +
  geom_hline(aes(yintercept = 0)) +
  labs(
    x = "Age Category",
    y = "Mean Difference Scores"
  )
```


## Random stuff

1.
Self-made discrete scale: scale_fill_manual(values = c("#F2FFDA", "#D3F3B4", "#BFE9B4", "#7FCD97", "#7FCDBB", "#41C6BD", "#41B6C4", "#1D9FC0", "#1D91C0", "#2273A9", "#225EA8", "#415EAC", "#253494", "#18337D", "#081D58", "#051644"))

2.
Things go on top of each other. If you make width of bar .5 and width of pattern in ggplot .4, you see things going out on the sides.

3.
How to make graphs APA style. See thesis project

4. change fig.height and width in r chunk code parameter to change the plot size. Looks like the default is 7.5 in height?

5. use str_wrap with width = 80 to make the title of a plot not go beyond the plot and be cut off. Creates a new line instead.


## To do list

- Add info about grammar of graphics
- Explore making ggplot APA style
  - Pattern library and the problem I ran into
- Explore different things one can do with ggplot
- Explore other libraries for data visualization too?
- Add in general commands for ggplot