---
title: "Data Wrangling of PAC Data Set"
---

## Pivoting from wide to long

I have a fairly tricky wide to long conversion. To get the right function, I will build it up in layers, starting with getting one and one thing correct and then adding additional challenges from the original data set until I get the entire conversion right.

```{r loading-libraries}
library(tidyverse)
```


First, I need to figure out how to move only part of the column name, but not all of it

Making some fake data:

```{r make-data-1}
id <- c(1001, 1002)
y2_YSR1 <- c(1, 3)
y3_YSR1 <- c(2, 4)
y2_YSR2 <- c(5, 6)
y3_YSR2 <- c(7, 8)

data <- tibble(id, y2_YSR1, y2_YSR2, y3_YSR1, y3_YSR2)

data
```

Now, I need to get the "yx" part as a separate column, keeping YSR1 and YSR2 as columns too

names_pattern = "^(yd)(.*)"

```{r pivot-longer-1}
data %>%
  pivot_longer(-id, names_to = c("Wave", ".value"), names_sep = "_")
```

I got this to work with names_sep, but not names_pattern. Trying again with names_pattern.

```{r make-data-2}
id <- c(1001, 1002)
y2YSR1 <- c(1, 3)
y3YSR1 <- c(2, 4)
y2YSR2 <- c(5, 6)
y3YSR2 <- c(7, 8)

data <- tibble(id, y2YSR1, y2YSR2, y3YSR1, y3YSR2)

data
```

```{r pivot-longer-2}
data %>%
  pivot_longer(-id, names_to = c("Wave", ".value"), names_pattern = "^(y\\d)(.*)")
```

Found the problem, my regex had only one backslash, when I needed two of them.
Now, the data looks like the way I want it. However, the data for wave one doesn't start with "y1", it is just blank. So I am thinking that the easiest way to work around it is to add y1 in front of all of those variables.
