---
title: Grading Contract
date: "`r format(Sys.time(), '%B %d %Y')`"
output: github_document
params:
  mondayweekone: "2025-01-13"
  thesisdeadline: "2025-03-21"
  springbreak: "2025-03-10"
  name: Linn Zapffe
  labs: 11
  portfolios: 10
  grade:
    label: Grade
    value: A
    input: select
    choices:
    - A
    - B
    - Pass
  deadline_day:
    label: Weekly Due Date
    value: Monday
    input: select
    choices:
    - Monday
    - Tuesday
    - Wednesday
    - Thursday
    - Friday
    - Saturday
    - Sunday
editor_options: 
  markdown: 
    wrap: 72
---

```{r setting number of assignments, include = FALSE}
# Setting the required number of labs and portfolio projects given the grade the student wants to make the contract for
if (params$grade == "A") {
  n_labs <- 11
  n_portfolios <- 10
} else if (params$grade == "B") {
  n_labs <- 8
  n_portfolios <- 6
} else if (params$grade == "Pass") {
  n_labs <- 8
  n_portfolios <- 4
} 
```

```{r defining variables, include = FALSE}
# Loading tidyverse, to deal with dates better
library(tidyverse)

# Defining the parameters as variables which makes it a little less cluttered during the contract itself
monday_week_one <- as.Date(params$mondayweekone)
deadline_day <- params$deadline_day
spring_break <- as.Date(params$springbreak)

# Defining a list with all the modules in it, which can be chosen from when making the schedule
modules_list = c(
  "Data Science",
  "Data and Visualization",
  "Tidyverse",
  "Data Diving with Tidyverse",
  "Effective Data Visualization",
  "Web scraping",
  "Functions and Automation",
  "Data and Ethics",
  "Modeling the Tidy way!",
  "Overfitting and Cross-validation",
  "Quantifying Uncertainty",
  "Base R and Simulating Data",
  "Large Language Models",
  "Interactive Web Apps",
  "Reproducible Reports",
  "Machine, Learn")

```

```{r dates and strings, include = FALSE}

# Defining a function we can call when we need a string in the contract with the modules, dates, and/or date for the deadlines
  # The function takes four parameters. The first is the date of the first Monday of classes, the second is the week number, the third is the module for that day, and the last one is the deadline day chosen earlier
string_details <- function(monday_week_one,
                    week_num,
                    module_num = NULL,
                    modules_list =  NULL,
                    deadline_day = NULL) {
  # Calculating the date for the weekday, based on the inputted week
  # For Monday only, as we use that to check whether it is the week of spring break, we keep a variable, date_monday, in date form. The as.Date() allows for the calculations to occur withouth treating any of the variables as string, won't run if it doesn't have the as.Date() function there.
  date_monday <- as.Date(monday_week_one + 7 * (week_num - 1))
  # Then we format the date into a string (happens automatically with format()) in the form of two numbers for day and month, e.g. "03/03"
    # For the rest of the days, we just keep this string version
  date_monday_str <- format(date_monday, format = "%m/%d")
  date_tuesday_str <- as.Date(monday_week_one + 1 + 7 * (week_num - 1)) %>%
    format(format = "%m/%d")
  date_wednesday_str_str <- as.Date(monday_week_one + 2 + 7 * (week_num - 1)) %>%
    format(format = "%m/%d")
  date_thursday_str <- as.Date(monday_week_one + 3 + 7 * (week_num - 1)) %>%
    format(format = "%m/%d")
  date_friday_str <- as.Date(monday_week_one + 4 + 7 * (week_num - 1)) %>%
    format(format = "%m/%d")
  date_saturday_str <- as.Date(monday_week_one + 5 + 7 * (week_num - 1)) %>%
    format(format = "%m/%d")
  date_sunday_str <- as.Date(monday_week_one + 6 + 7 * (week_num - 1)) %>%
    format(format = "%m/%d")
  # Format the week number with two numbers, e.g. "02"
  module_string <- sprintf("%02d", module_num)
  
  # Creating the string, based on what type of information is given. This determines what format the string should have
  # First if checks for whether it is spring break. If it is, then there is no module to do
  if (date_monday == spring_break) {
    # paste0 adds the parts in it together to a string, separated by space
    str_line <- paste0("Spring break! :", date_monday_str, " - ", date_sunday_str)
  } else if (!is.null(deadline_day)) {
    if (deadline_day == "Monday") {
      str_line <- date_monday_str
    } else if (deadline_day == "Tuesday") {
      str_line <- date_tuesday_str
    } else if (deadline_day == "Wednesday") {
      str_line <- date_wednesday_str
    } else if (deadline_day == "Thursday") {
      str_line <- date_thursday_str
    } else if (deadline_day == "Friday") {
      str_line <- date_friday_str
    } else if (deadline_day == "Saturday") {
      str_line <- date_saturday_str
    } else if (deadline_day == "Sunday") {
      str_line <- date_sunday_str
    }
    # Creating a string with the module, module number, dates for the week, and the topic for the module
  } else{
    str_line <- paste0("Module ", module_num, ", ", date_monday_str, " - ", date_sunday_str, " : ", modules_list[module_num])
  }
  # Return the string so that it can be accessed outside of the function
  return(str_line)
}

```

<!-- This contract is adapted from Annie Somerville's contract https://github.com/anniehsom -->

# Contract Grades – “`r params$grade`” Contract

I, `r params$name`, agree to :

1)  Earn “Satisfactory” on `r n_labs` self-paced lab assignments on the
    schedule specified below. Getting a grade of "Satisfactory" requires
    that:

-   I have attempted all steps of the assignment. If unable to complete
    steps due to unresolved questions, I have asked questions in Github
    FAQs or during the help sessions, or I have Googled the issue and
    shown effort to resolve questions;
    -   I have checked that files, repos, or projects appear on my
        Github account and are accessible;
    -   I feel like I have spent an adequate amount of time and energy
        on assignments;

2)  Produce at least `r n_portfolios` “Satisfactory” portfolio pieces on
    the schedule specified below. Getting a grade of "Satisfactory"
    requires that:
    -   I have designed a portfolio project that I think challenges my
        understanding of programming, explores a modules_list I want to
        know more about, or targets something I feel like I need more
        practice with;
    -   I have checked that files, repos, or projects appear on my
        Github account and are accessible;
    -   I feel like I have spent an adequate amount of time and energy
        on the projects;

## SCHEDULE


- `r string_details(monday_week_one, 1, 1, modules_list)`
   - Finish Lab 1 by `r deadline_day`:
        `r string_details(monday_week_one, 2, deadline_day=deadline_day)`

- `r string_details(monday_week_one, 2, 2, modules_list)`
    - Finish Lab 2 by `r deadline_day`:
        `r string_details(monday_week_one, 3, deadline_day=deadline_day)`
    - Propose your contract to Mason by `r deadline_day`:
        `r string_details(monday_week_one, 3, deadline_day=deadline_day)`
- `r string_details(monday_week_one, 3, 3, modules_list)`
    - Finish Lab 3 by `r deadline_day`:
        `r string_details(monday_week_one, 4, deadline_day=deadline_day)`
    - Finish portfolio piece 1 by `r deadline_day`:
        `r string_details(monday_week_one, 4, deadline_day=deadline_day)`
- `r string_details(monday_week_one, 4, 4, modules_list)`
    - Finish Lab 4 by `r deadline_day`:
        `r string_details(monday_week_one, 5, deadline_day=deadline_day)`
- `r string_details(monday_week_one, 5, 5, modules_list)`
    - Finish Lab 5 by `r deadline_day`:
        `r string_details(monday_week_one, 6, deadline_day=deadline_day)`
    - Finish portfolio piece 2 by `r deadline_day`:
        `r string_details(monday_week_one, 6, deadline_day=deadline_day)`
- `r string_details(monday_week_one, 6, 6, modules_list)`
    - Finish Lab 6 by `r deadline_day`:
        `r string_details(monday_week_one, 7, deadline_day=deadline_day)`
    - Schedule a check-in with Mason by `r deadline_day`:
        `r string_details(monday_week_one, 7, deadline_day=deadline_day)`
- `r string_details(monday_week_one, 7, 7, modules_list)`
    - Finish Lab 7 by `r deadline_day`:
        `r string_details(monday_week_one, 8, deadline_day=deadline_day)`
    - Finish portfolio piece 3 by `r deadline_day`:
        `r string_details(monday_week_one, 8, deadline_day=deadline_day)`
- `r string_details(monday_week_one, 8, 8, modules_list)`
    - Finish Lab 8 by `r deadline_day`:
        `r string_details(monday_week_one, 9, deadline_day=deadline_day)`
- `r string_details(monday_week_one, 9, 9, modules_list)`
    - Finish Lab 9 by `r deadline_day`:
        `r string_details(monday_week_one, 10, deadline_day=deadline_day)`
    - Finish portfolio piece 4 by `r deadline_day`:
        `r string_details(monday_week_one, 10, deadline_day=deadline_day)`
- `r string_details(monday_week_one, 10, 10, modules_list)`
    - Finish Lab 10 by `r deadline_day`:
        `r string_details(monday_week_one, 11, deadline_day=deadline_day)`
- `r string_details(monday_week_one, 11, 11, modules_list)`
    -`r if(params$labs>=11) paste0(" Finish Lab 11 by ",deadline_day,": ",string_details(monday_week_one, 12, deadline_day=deadline_day))`
    - Finish portfolio piece 5 by `r deadline_day`:
        `r string_details(monday_week_one, 12, deadline_day=deadline_day)`
- `r string_details(monday_week_one, 12, 12, modules_list)`
    `r if(params$labs>=12) paste0(" - Finish Lab 12 by ",deadline_day,": ",string_details(monday_week_one, 12, deadline_day=deadline_day))`
    - Finish portfolio piece 6 by `r deadline_day`:
        `r string_details(monday_week_one, 13, deadline_day=deadline_day)`
- `r string_details(monday_week_one, 13, 13, modules_list)`
    `r if(params$labs>=13) paste0(" - Finish Lab 13 by ",deadline_day,": ",string_details(monday_week_one, 14, deadline_day=deadline_day))`
    - Finish portfolio pieces 7 and 8 by `r deadline_day`:
        `r string_details(monday_week_one, 14, deadline_day=deadline_day)`
- `r string_details(monday_week_one, 14, 14, modules_list)`
    - Finish portfolio piece 9 by `r deadline_day`:
        `r string_details(monday_week_one, 15, deadline_day=deadline_day)`
- `r string_details(monday_week_one, 15, 15, modules_list)`
    - Finish portfolio piece 10 by `r deadline_day`:
        `r string_details(monday_week_one, 16, deadline_day=deadline_day)`
- Portfolio completion
    - Due on `r deadline_day`:
        `r string_details(monday_week_one, 16, deadline_day=deadline_day)`

## Further guidelines

-   According to reasons discussed and agreed upon in advance with the
    professor, there are some situations which might require some of
    these deadlines to be moved. If this is the case, I will reach out
    to the professor and agree on another deadline. The number of
    assignments that have to be completed satisfactory will not change.
-   If the professor has any concerns about the assignments and whether
    they reach the standards of being completed "satisfactory", the
    professor will communicate these concerns with me and offer an
    opportunity to correct them.
-   I pertain the right to switch or switch out any of the modules that
    do not contain a lab with another module in the course. This can be
    done in advance or while working on a module.
-   Any part of this contract can be changed before the final grade is
    set if agreed upon by both myself and the professor.

Change out any labs too (there are 13)

Test string "`r string_details(monday_week_one, 2, 2, modules_list)`" 

## SCHEDULE


- [ ] `r string_details(monday_week_one, 1, 1, modules_list)`
  - Finish Lab 1 by `r deadline_day`:
        `r string_details(monday_week_one, 2, deadline_day=deadline_day)`
- `r string_details(monday_week_one, 1, 1, modules_list)`
   - Finish Lab 1 by `r deadline_day`:
        `r string_details(monday_week_one, 2, deadline_day=deadline_day)`
        
        
        
- [ ] `r string_details(monday_week_one, 2, 2, modules_list)`
    - [ ] Finish Lab 2 by `r deadline_day`:
        `r string_details(monday_week_one, 3, deadline_day=deadline_day)`
    - [ ] Propose your contract to Mason by `r deadline_day`:
        `r string_details(monday_week_one, 3, deadline_day=deadline_day)`
- [ ] `r string_details(monday_week_one, 3, 3, modules_list)`
    - [ ] Finish Lab 3 by `r deadline_day`:
        `r string_details(monday_week_one, 4, deadline_day=deadline_day)`
    - [ ] Finish portfolio piece 1 by `r deadline_day`:
        `r string_details(monday_week_one, 4, deadline_day=deadline_day)`
- [ ] `r string_details(monday_week_one, 4, 4, modules_list)`
    - [ ] Finish Lab 4 by `r deadline_day`:
        `r string_details(monday_week_one, 5, deadline_day=deadline_day)`

